???A quantitative framework for investigating the reliability of network construction

### 2017-11-18 manuscript/CI_interactions.tex
Gravel first read

    % Present Salix data with holes as an example of uncertainty in data. Focus just on the white and grey, ignore uncertainty in the blacks
    In modern community ecology, species assemblages are increasingly described and analyzed as networks~\citep{Thompson2012,RecentReviews}. Such representations offer convenient summaries of how communities are constructed since they simultaneously account for species composition and their interactions. Species are represented as nodes, hence a tabulation of the nodes present and of their relative abundances form the basis for traditional metrics of community composition. On top of this information, interactions are represented by links, so that networks provide additional, higher-order information on community structure. Metrics of how networks are structured at the level of links emerging from individual species (such as link density, vulnerability and degree distribution), and of network-level structure (such as nestedness, connectance and modularity), then offer key descriptors of how species interactions are structured across the globe~\citep{Schleuning2012,Morris2014,XX}, valuable proxies of community dynamics~\citep{Thebault2010,XX} and [many other nice things]. Thus, deriving robust descriptors of network properties is a key objective for modern community ecology.

    Importantly, empirical descriptions of ecological networks are usually considered to be static representations of the communities and interactions they describe. That is, whether the network is assembled based on aggregated data, a single intensive ``snapshot" sample, or expert knowledge, interactions are assumed to occur at the same frequencies wherever and whenever the network is observed~\citep{Olesen2011a}. This assumption conflicts with known variation in interactions over time~\citep{Kitching1987,Olesen2011a} and space~\citep{Kitching1987,Baiser2012} and between individuals of a given species~\citep{Pires2011a,Fodrie2015,Novak2015}. Taking the variation in interaction frequencies into account would therefore represent a substantial increase in the realism of ecological networks as representations of ecological communities.

    Efforts to account for variance in interaction frequencies are complicated by the fact that not only interaction frequencies but also network structure vary over space and time. Community composition and species abundances vary from site to site~\citep{Baiser2012} and over time within a site~\citep{Olesen2011a}. The removal of a species from a site will obviously also remove its interactions. Interactions can also be lost, however, if participating species remain present but do not co-occur temporally or are too rare to detect each other~\citep{Tylianakis2010}, or through changes to individual preferences~\citep{Fodrie2015}.

    Beyond ``true" variation in network structure, several researchers have pointed to the importance of sampling intensity for the assessment of network structure (e.g.,~\citealp{Martinez1999,Bluthgen2006,XX}). Despite this, the acknowledgment that insufficient sampling effort limits our ability to describe variation in ecological networks has not yet led to a quantitative framework to deal with the uncertainty generated by spatiotemporal variation in ecological interactions and by sampling.

    In this study, we develop analytical tools to capture the uncertainty in the estimation of pairwise interaction probability. To define the problem, we first discuss the pervasive implications of uncertainty in the estimation of pairwise interactions on the properties of ecological networks. We then define a framework which identifies and dissects the different contributions to uncertainty of interaction. We illustrate this framework with a simple quantitative example. Next, we use the framework to provide analytical criteria for improving the sampling of rare interactions. Finally, we offer tangible recommendations for improved descriptors of ecological interactions. Through these efforts, we demonstrate both the utility of our approach and the importance of acknowledging the uncertainty inherent in ecological networks.

@@

  To define the problem associated with describing ecological interaction networks, we will start from the perspective of an empirical community ecologist faced with the task of describing an interaction network. This ecologist will be interested in generating a description of the blocks present, and the links between them (Roslin and Majaneva 2017). In other words, he or she seeks to fill in a matrix of species by species interactions. Importantly, the information sought is conveyed by both the presence and \emph{absence} of links. An observed link will always remain an observed link, but as to why given links are not observed, there may be multiple contributions. Thus, we will set off by contemplating what factors may contribute to links missing from the matrix. We consider that there are three nested levels of uncertainty making the occurrence of any interaction a stochastic process:

@@

    First, and most fundamentally, we do not know whether or not a pair of species have the appropriate characteristics to interact. We define the probability of an interaction $L$ as $P(L)=\lambda$. Obviously, because $k$ (the number of observed interactions) is 0, it is possible that the two species cannot interact even if there were no time or environmental constraints preventing the interaction. 
    % [[The previous sentence doesn't follow, k hasn't been defined and we don't know that it's zero. Also, are we assuming that these interactions have been observed over a large area/time like the Kopelke data? Otherwise couldn't something about that particular site prevent species interacting even if they otherwise could?]] 
    % [[With defining k I think this is fine... site uncertainty will come in later, and it's not necessary to assume anything about the sampling design to get from not observing two species to interact to suspecting that they don't (degree of confidence depending on the species involved). I've toned down "very likely" to "possible" though - does that help?]]
    But it is nonetheless possible that the interaction is a rare phenomenon that has never been documented before. This source of uncertainty is the one documented by trait-matching models~\citep{}. It arises because every model is imperfect and lacks information (i.e. traits) about the constraints on the interaction. In other words, with sufficient sampling and all information accessible, this interaction probability should either tend to 0 or to 1 and the uncertainty vanishes. 
    % [[For species which actually co occur (?)]]
    % [[I think that gets covered under "sufficient sampling"]]

@@

    It could be that the interaction is feasible, i.e. $L=1$, but that it does not occur at a given location or moment in time because there are local constraints, such as inclement weather or lack of suitable habitat, that prevent the interaction from occurring. We define the realisation of the interaction process $X$, given that the interaction is feasible, as a stochastic process with associated probability $P(X|L)=\chi$. This phenomenon of interaction contingencies is usually not considered in network studies, but there is a rich literature in community ecology about the contingencies of interactions~\citep{}. 
    % We imagine for instance that an interaction between a gall and a parasitoid might not be recorded at a location because it was the wrong time of the year and the gall had not yet formed (phenological constraints), a late frost killed the larvae (an abiotic environmental constraint), another parasitoid species competitively excluded the parasitoid species of interest (a biotic environmental constraint) or simply the parasitoid is too rare to interact with all of its viable hosts (an abundance constraint).

@@

    Lastly, measurement errors are always a source of uncertainty in the observation of ecological processes. We define the detection of an interaction $D$, given an interaction is feasible and occurs in the local conditions, as a stochastic process with associated probability $P(D|X,L)=\delta$. Detection failure could happen for several reasons, for instance if the rearing of a parasitoid fails with inappropriate lab conditions, because of species mis-identification, or because the interaction is very rare. Some sources of detection error could be minimised with appropriate sampling effort ($\delta$ will converge to one with increasing number of samples), but other sources are often difficult to reduce (e.g. the occurrence of cryptic species might require molecular analysis for appropriate taxonomic identification).

@@

    The combination of these three sources of uncertainty together leads to many potential explanations for the observation of an absence of interaction, but only the situation where $L = 0$ (a true absence) is relevant for the investigation of networks. While the observation of an interaction is straightforward to interpret, absence of interactions must be decomposed into different quantities. It is particularly important to rule out the situations where $D=0 \cup X = 1 \cup L=1$, i.e. the interaction occurred at the location but was not observed, and $D=1 \cup X = 0 \cup L =1$, the interaction is feasible and would have been detected but did not occur at the local site.

  It is important to keep in mind that for network analysis, ecologists seek to measure the occurrence of true absences, which is the joint event $L=0 \cup X=1 \cup D=1$), but in reality they measure the marginal probability $P(L) = k/n$ where $k$ is again the number of observed interactions and $n$ the number of observed co-occurances. % I'm pretty sure this should be L=0 but X=1 and D=1: interaction not feasible but site and detection perfect.

@@

  The basic information that we are collecting consists of matrices of interaction partners. Our count of interactions between species A and species B reflects uncertainty with respect to whether two species are capable of interacting when co-occurring at a given site, chance variation in whether the interaction is realised,  and uncertainty with respect to whether it is detected when realised. In this case, what can we do to pin down any of these different types of variation?

  A simple example will illustrate the problem: assume that we want to infer the probability of an interaction between two species, A and B. Now assume that in reality, interaction between A and B is completely impossible (i.e. the true $\lambda$= 0). Given no added detection error in observing the incidence of the interaction, then a single observation of species co-occurrence will reveal (very) little regarding the probability of the interaction – the confidence limits for one observation essentially spans from 0 to 1 (Fig.~\ref{upper_limits}). For 5 observations, this confidence limit still extends to 0.5. Only with thirty observations will it shrink to 0.1. Thus, adding (many) more observations is certainly useful in controlling uncertainty due to stochasticity inherent in the process itself.

@@

  As a solution towards deciding whether unobserved interactions were undetected due to sampling, or whether they truly do not occur,~\citet{Weinstein2017} used daily repeated sampling rounds to estimate the daily probability of detecting a hummingbird interaction, and to thereby model the observation and process mechanisms~\citep{Weinstein2017}. While conceptually attractive, this approach is unsuitable for interactions occurring over longer time scales (e.g., associations between hosts and parasitoids with a single generation per year), or very rare interactions (since no matter how frequently and extensively we sample, there will still be some interactions lurking around). What is worse, the problem persists that if a given interaction is not observed on a given day, this could be either because it was impossible that day despite being otherwise feasible (P($X|L=1$=0), because interaction between the two did occur but could not be observed (P($D|X$=1,L=1)=0), or from any combination of the two. From a conceptual perspective, this approach will thus fail to satisfactorily distinguish between sources of uncertainty. An interaction not observed on day $t$ may have gone undetected either because there were constraints on the process (i.e. the probability $\lambda$ was really zero) or because of chance variation in the realised incidence of an event associated with a finite probability. Most importantly, if two species are never observed co-occurring during several days of sampling then we have learned nothing about their probability of interacting if they should ever co-occur. In other words, there is no information without co-occurrence.

@@

@@

  To progressively dissect the different contributions to uncertainty, we will start by considering how we could naively quantify interaction probability and its associated uncertainty \emph{for an interaction that has not yet been observed}. We consider that a pair of species have been observed $n$ times co-occurring at one location, and the number of times they have been observed interacting is $k = 0$. We now aim to evaluate the uncertainty of this interaction. The occurrence of an interaction is a Bernoulli trial. Consequently, the number of success[es] $k$ over $n$ trials will follow a binomial distribution:

@@

  The situation where $k = 0$ is also problematic because we have very little information with which to estimate $\lambda$, in particular when $n$ is also small. We therefore propose a Bayesian approach to solve this problem, building on two relevant features of such statistics: the posterior distribution of parameter estimates and usage of prior information. Before moving to this direction, we will first take time to better describe what is exactly meant by an interaction probability and appropriately define sources of uncertainty.

  So many sources of interaction pooled into a single entity; hard to estimate.

  MLE is zero if no interaction is detected, but there may still be a substantial probability of interaction should the two species co-occur. [This is particularly important in the context of a changing world, where species not co-occurring today – or species not detected co-occurring today – may well do so tomorrow, forming complex interactions.]

@@

      We here adopt a Bayesian approach to estimate the distribution of the parameter $\theta$, the maximum likelihood estimate of the interaction probability. According to Bayes principle, the posterior distribution of $\theta$ is:

      \begin{equation}
        \underbrace{P(\theta|X,n)}_{Posterior} = \frac{\overbrace{P(X|\theta,n)}^{Likelihood}\overbrace{P(\theta)}^{Prior}}{\underbrace{P(X|N)}_{Normaliser}} .
        \label{posterior}
      \end{equation}

      According to the above description, the likelihood is simply the binomial distribution (Eq.~\ref{likelihood}). Since $\theta$ is a probability it is bounded between 0 and 1 and consequently the most appropriate prior distribution is the beta:

      \begin{equation}
        \theta \sim Beta(\alpha,\beta) , \label{prior}
      \end{equation}

@@

     In some cases it can be complicated to compute the normaliser, but fortunately in our case an analytical solution exists. The beta-binomial distribution is a conjugate distribution of the binomial distribution. This allows us to analytically compute the posterior distribution of a binomial model with a beta prior distribution. We can re-write the posterior distribution of $\theta$ as:

      \begin{equation}
        P(\theta|k,n) = \frac{\theta^{\alpha+k-1}(1-\theta)^{\beta+n-k-1}}{B(\alpha+k,\beta+n-k)} , \label{posterior}
      \end{equation}

      \noindent where the function $B$ is the beta function:

      \begin{equation}
        Beta(\alpha+k,\beta+n-k) = \frac{\Gamma(\alpha+k)\Gamma(\beta+n-k)}{\Gamma(\alpha+\beta+n)} . \label{betafunction}
      \end{equation}

      The posterior distribution of $\theta$ therefore follows the beta distribution with new parameters $\alpha'= \alpha+k$ and $\beta'=\beta+n-k$. The weight of the prior on the posterior distribution is understood from these definition of the parameters: the difference between the posterior and the prior will increase with $k$ and $n-k$. When plotted, we find the shape of the distribution gets narrower with $k$ and $n$ (Fig.~\ref{Salix_pdfs}). 
      %
      %  The posterior distribution of $\theta$ could be computed with R with the following command:
      %
      %
      %  \vspace{12pt}
      %  \noindent\emph{dbeta(x = theta, shape1 = alpha+k, shape1 = beta+n-k)}
      %  \vspace{12pt}

      \begin{figure}[ht]
        \label{Salix_pdfs}
        \caption{Using prior distribution based on the \emph{Salix}-galler and galler-parasitoid networks in~\citet{Barbour2016}, we can calculate posterior distributions for the probability of interaction ($\lambda$) between two species that have not yet been observed interacting. Here we show posterior distributions for each $\lambda$ in each network ranging from the prior distribution (N=0 observed co-occurrence) to the distribution obtained when the pair of species has been observed co-occurring 100 times. The distribution narrows and approaches zero as the sample size increases. \textbf{A)} The posterior distributions for the \emph{Salix}-galler component are always wider and farther from zero than those for \textbf{B)} the galler-parasitoid component. This is likely because the prior distribution for the galler-parasitoid component of the network was both narrower and had a mean closer to that in the~\citet{Kopelke2017} data than the \emph{Salix}-galler component.}
        \includegraphics*[width=.8\textwidth]{figures/Salix_Galler_pdfs_increasing_N.eps}
        \end{figure}

@@

      The fact that the posterior distribution of $\theta$ follows a beta distribution makes it straightforward to compute moments and other properties.

      The \textbf{average} of $\theta$ is: 
          \begin{equation}
            \bar{\theta} = \frac{\alpha+k}{\alpha+\beta+n} ,
            \label{mean}
          \end{equation}

        and its \textbf{variance} is:  
          \begin{equation}
            Var(\theta|k) = \frac{(\alpha + k)(\beta + n - k)}{(\alpha + \beta + n)^{2}(\alpha + \beta + n +1)}
            \label{variance}
          \end{equation}

        The \textbf{mode} of the distribution is:
          \begin{equation}
            \hat{\theta} = \frac{\alpha + k - 1}{\alpha + \beta + n - 2} .
            \label{mode}
          \end{equation}

@@

      Parameters $\alpha$ and $\beta$ determine the shape of the prior distribution, which follows a beta distribution. These are called hyper parameters. Below we identify four ways to formulate the prior distribution of $\theta$.

@@

          In absence of any external information, an uninformative prior is the most conservative hypothesis for the distribution of $\theta$. The beta distribution is in this case a uniform distribution  for $Beta(\alpha=1,\beta=1)$.

@@

          [[Should we still have this? At the workshop we talked about how interaction probabilities depend on the degrees of both species and so I'm using predicted interaction probabilities (degreeA*degreeB) in the galler example. A writeup for that process follows the original text:]]
          The degree of a node in a network is defined as the number of interactions it has to other nodes. The degree distribution of a network is then the probability distribution of these degrees over the whole network. Connectance is an average property for a complete given network, and it represents the expected degree (expected number of interactions per species) standardised by the number of species in the network. The standardised degree could therefore be interpreted as an interaction probability. It is consequently possible to use the degree distribution to inform the prior distribution. The degree distribution could come from several networks, from a similar network (e.g. a known network at slightly different location) or from the network of interest if interaction probabilities for some species are already documented. The latter approach allows us to apply information from known, abundant species to the rarest species for which interactions are less frequently documented. The procedure for the estimation of the hyper parameters follows the same approach as described above for connectance except that each measurement is at the individual species level instead of the network level.

          If our focal network describes a similar system to that in a known network, we can use the distribution of interaction probabilities in that network to inform our prior. The probability of any interaction $L_ij$ depends on the degrees of species $i$ and $j$. The degree of a node is defined as the number of interactions it has with other nodes. Connectance, as the average number of links per species in the whole network, is thus also the expected degree for any given species normalised by the number of species in the network. Using normalised degrees $\Delta_i$ and $\Delta_j$ (i.e., degrees divided by the number of species in the network), we can obtain the probability of interaction $L_ij$=$\Delta_i\times\Delta_j$. The distribution of these interaction probabilities derived from a similar network or a suite of networks can be used to establish a prior distribution before any data from the focal network are collected. We could also use this approach to build a prior distribution using known interaction probabilities for some species in the focal network. The latter approach allows us to apply information from known, abundant species to the rarest species for which interactions are less frequently documented. The procedure for the estimation of the hyper parameters follows the same approach as described above for connectance except that each measurement is at the individual interaction level instead of the network level.

@@

          [Tomas adds: This is a bit of a stump]
          It is possible to estimate the probability of interaction between a pair of species if there is knowledge of their traits and some functions relating trait matching to interaction probability~\citep{Morales-Castilla2015}. There are several techniques available to perform this inference of interaction probability~\citep{} (REFS). Note that in this case the prior might not be beta distributed and numerical methods might be required to compute the posterior distribution.

>>>>

    % Present Salix data with holes as an example of uncertainty in data. Focus just on the white and grey, ignore uncertainty in the blacks

  %DG : I cut down significantly the first paragraph, as I think we could assume readers know about networks and don't need to be convinced of their usage. 
   Representation of species assemblages as a network offers a convenient summary of how communities are constructed, as it simultaneously account for species composition and their interactions. A tabulation of the nodes and of their relative abundances form the basis for traditional metrics of community composition. On top of this information, interactions are represented by links, so that networks provide additional, higher-order information on community structure. Empirical descriptions of ecological networks are however usually considered to be static representations of the communities and interactions they describe. This approach contrasts significantly with the widely recognized stochasticity of ecological communities \citep{Gotelli2000}. That is, whether the network is assembled based on aggregated data, a single intensive ``snapshot" sample, or expert knowledge, interactions are assumed to occur deterministically wherever and whenever the community is observed~\citep{Olesen2011a}. This assumption conflicts with known variation in interactions over time~\citep{Kitching1987,Olesen2011a} and space~\citep{Kitching1987,Baiser2012} and between individuals of a given species~\citep{Pires2011a,Fodrie2015,Novak2015}.

    Efforts to account for the variability in interaction occurrence and frequency are complicated by the fact that not only species but also interactions vary over space and time. Community composition and species abundances vary from site to site~\citep{Baiser2012} and over time within a site~\citep{Olesen2011a}. The removal of a species from a site will obviously also remove its interactions, but the co-occurrence of potentially interacting species do not guarantee they will. Interactions can be lost, or not detected, if participating species remain present but do not co-occur temporally, are too rare to detect each other~\citep{Tylianakis2010}, because of environmental contingencies \citep{Poisot2015a} or through changes to individual preferences~\citep{Fodrie2015}.

    Beyond ``true" variation in network structure, several researchers have pointed to the importance of sampling intensity for the assessment of network structure (e.g.,~\citealp{Martinez1999,Bluthgen2006,XX}). Accumulation of interactions with increasing sampling effort suggests that it is even more challenging to document them than species richness \citep{Poisot2012}. As a result, it was proposed that interactions should be described probabilistically and network metrics computed accordingly \citep{Poisot2015b}. Despite this, the acknowledgment that insufficient sampling effort limits our ability to describe variation in ecological networks has not yet led to a quantitative methodology to deal with the uncertainty generated by spatiotemporal variation in ecological interactions and by sampling.

% DG : I wonder if a paragraph summarizing the extent of the problem with the salix data should not be provided here. Either within the text, or a box. But we need to convingly show how large is this problem even for such an extensive dataset

    In this study, we develop analytical tools to capture the uncertainty in the estimation of pairwise interactions. We focus on binary interactions as a first step, but the framework could be expanded to deal with interaction frequencies and strength. To define the problem, we first discuss the pervasive implications of uncertainty in the estimation of pairwise interactions on the properties of ecological networks. We then define a framework which identifies and dissects the different contributions to uncertainty of interaction. We illustrate this framework with a simple quantitative example using the most extensively sample network to date. Next, we use the framework to provide analytical criteria for improving the sampling of rare interactions. Finally, we offer tangible recommendations for improved descriptors of ecological interactions. Through these efforts, we demonstrate both the utility of our approach and the importance of acknowledging the uncertainty inherent in network studies.

@@

  To define the problem associated with describing ecological interaction networks, we will start from the perspective of an empirical community ecologist faced with the task of describing an interaction network. This ecologist will be interested in generating a description of the blocks present, and the links between them (Roslin and Majaneva 2017).  Importantly, the information sought is conveyed by both the presence and \emph{absence} of links. An observed link will always remain an observed link, but as to why given links are not observed, there may be multiple contributions. Thus, we will set off by contemplating what factors may contribute to links missing from the matrix. We consider that there are three nested levels of uncertainty making the occurrence of any interaction a stochastic process:

@@

    First, and most fundamentally, we do not know whether or not a pair of species have the appropriate characteristics (or traits) to interact. We define the probability of an interaction $L$ given those characteristics $\mathbf{T}$ as $P(L | \mathbf{T})=\lambda$. Obviously, if $k$ (the number of observed interactions) is 0, it is possible that the two species cannot interact even if there were no time or environmental constraints preventing it to occur. In such case, $\lambda$ would take a value of 0. 
    % [[The previous sentence doesn't follow, k hasn't been defined and we don't know that it's zero. Also, are we assuming that these interactions have been observed over a large area/time like the Kopelke data? Otherwise couldn't something about that particular site prevent species interacting even if they otherwise could?]] 
    % [[With defining k I think this is fine... site uncertainty will come in later, and it's not necessary to assume anything about the sampling design to get from not observing two species to interact to suspecting that they don't (degree of confidence depending on the species involved). I've toned down "very likely" to "possible" though - does that help?]]
    But it is nonetheless possible that the interaction is a rare phenomenon that has never been documented before, with $\lambda>0$. This source of uncertainty is the one documented by trait-matching models~\citep{Bartomeus2016}. It arises because every model is imperfect and lacks information (i.e. traits) about the constraints on the interaction. In other words, with sufficient sampling and all information accessible, this interaction probability $\lambda$ should either tend to 0 or to 1 and the uncertainty vanishes. 
    % [[For species which actually co occur (?)]]
    % [[I think that gets covered under "sufficient sampling"]]

@@

    It could be that we know an interaction is feasible, i.e. $L=1$, but that it does not occur at a given location or moment in time because there are local constraints, such as inclement weather or lack of suitable habitat, that prevent the interaction from occurring. We define the realisation of the interaction process with the variable $X$, given that the interaction is feasible, as a stochastic process with associated probability $P(X|L)=\chi$. This phenomenon of interaction contingencies is usually not considered in network studies, but there is a rich literature in community ecology about the contingencies of interactions~\citep{}. 
    % We imagine for instance that an interaction between a gall and a parasitoid might not be recorded at a location because it was the wrong time of the year and the gall had not yet formed (phenological constraints), a late frost killed the larvae (an abiotic environmental constraint), another parasitoid species competitively excluded the parasitoid species of interest (a biotic environmental constraint) or simply the parasitoid is too rare to interact with all of its viable hosts (an abundance constraint).

@@

    Lastly, measurement errors are always a source of uncertainty in the observation of ecological processes. We define the detection of an interaction with the variable $D$, given an interaction is feasible and occurs in the local conditions, as a stochastic process with associated probability $P(D|X,L)=\delta$. Detection failure could happen for several reasons, for instance if the rearing of a parasitoid fails with inappropriate lab conditions, because of species mis-identification, or because the interaction is very rare. Some sources of detection error could be minimised with appropriate sampling effort ($\delta$ will converge to one with increasing number of samples), but other sources are often difficult to reduce (e.g. the occurrence of cryptic species might require molecular analysis for appropriate taxonomic identification).

@@

    The combination of these three sources of uncertainty together leads to many potential explanations for the observation of an absence of interaction, but only the situation where $L = 0$ (a true absence) is relevant for the investigation of networks. While the observation of an interaction is straightforward to interpret, absence of interactions must be decomposed into different quantities. It is particularly important to rule out the situations where $D=0 \cup X = 1 \cup L=1$, i.e. the interaction occurred at the location but was not observed, and $D=1 \cup X = 0 \cup L =1$, the interaction is feasible and would have been detected but did not occur at the local site. Ecologists seek to measure the occurrence of true absences, which is the joint event $L=0 \cup X=1 \cup D=1$), but in reality they measure the marginal probability $P(L) = k/n$ where $k$ is again the number of observed interactions and $n$ the number of observed co-occurrences. % I'm pretty sure this should be L=0 but X=1 and D=1: interaction not feasible but site and detection perfect.
    %DG: YES

@@

  The basic information that we are collecting consists of matrices of interaction partners. Our count of interactions between two species reflects uncertainty with respect to whether two species are capable of interacting when co-occurring at a given site ($\lambda$), chance variation in whether the interaction is realised ($\chi$),  and uncertainty with respect to whether it is detected when realised ($\delta$). In this case, what can we do to pin down any of these different types of variation?

% DG: it's bizarre to present here the outcome of the naive binomial formulation, while the reasoning is provided later. My solution is to move the following section on the binomial model along with the figure within a box. It's not central, but formalizes the problem. It would respect the order of appearance and better emphasize the bayes section.

  A simple example will illustrate the problem (the full reasoning is provided below): assume that we want to infer the probability of an interaction between two species, A and B. Now assume that in reality, interaction between A and B is completely impossible (i.e. the true $\lambda=0$), but the observer does not know it and seeks to estimate this interaction probability. The record of the number of observed interactions will follow a binomial distribution, with number of trials $k$ and number of observations $n$, for which we can compute the confidence interval of the estimated probability $\lambda$. Given no added detection error in observing the incidence of the interaction, then a single observation of species co-occurrence will reveal (very) little regarding the probability of the interaction – the confidence limits for one observation essentially spans from 0 to 1 (Fig.~\ref{upper_limits} - using Clopper-Pearson's true confidence interval). For 5 observations, this confidence limit still extends to 0.5. Only with thirty observations will it shrink to 0.1. Thus, adding (many) more observations is certainly useful in controlling uncertainty due to stochasticity inherent in the process itself, but the number of observations nonetheless needs to be very high.

@@

  As a solution towards deciding whether unobserved interactions were undetected due to sampling, or whether they truly do not occur,~\citet{Weinstein2017} used repeated sampling rounds to estimate the daily probability of detecting a hummingbird interaction, and to thereby model the observation and process mechanisms. While conceptually attractive, this approach is unsuitable for interactions occurring over longer time scales (e.g., associations between hosts and parasitoids with a single generation per year), or very rare interactions (since no matter how frequently and extensively we sample, there will still be some interactions lurking around). What is worse, the problem persists that if a given interaction is not observed on a given day, this could be either because it was impossible that day despite being otherwise feasible (P($X|D=1,L=1$=0), because interaction between the two did occur but could not be observed (P($D|X$=1,L=1)=0), or from any combination of the two. From a conceptual perspective, this approach will thus fail to satisfactorily distinguish between sources of uncertainty. An interaction not observed on day $t$ may have gone undetected either because there were constraints on the process (i.e. the probability $\lambda$ was really zero) or because of chance variation in the realised incidence of an event associated with a finite probability. Most importantly, if two species are never observed co-occurring during several days of sampling then we have learned nothing about their probability of interacting if they should ever co-occur. In other words, there is no information without co-occurrence.

@@

% DG : I propose to maintain part of this paragraph that was commented out, just to keep the transition. 
The bottom line is that separating different sources of uncertainty is difficult indeed. As an alternative to abandoning empirical networks or continuing to ignore the uncertainty inherent in undetected observations, we propose that some insight regarding the detectability of interactions between species not found co-occurring may be gained from data on other species between which interactions are known to be feasible.

@@

  To progressively dissect the different contributions to uncertainty, we will start by considering how we could naively quantify interaction probability and its associated uncertainty \emph{for an interaction that has not yet been observed}. We consider that a pair of species have been observed $n$ times co-occurring at one location, and the number of times they have been observed interacting is $k = 0$. We now aim to evaluate the uncertainty of this interaction. The occurrence of an interaction is a Bernoulli trial. Consequently, the number of successes $k$ over $n$ trials will follow a binomial distribution:

@@

% DG: The following should be removed to avoid repetitions
%  The situation where $k = 0$ is also problematic because we have very little information with which to estimate $\lambda$, in particular when $n$ is also small. We therefore propose a Bayesian approach to solve this problem, building on two relevant features of such statistics: the posterior distribution of parameter estimates and usage of prior information. Before moving to this direction, we will first take time to better describe what is exactly meant by an interaction probability and appropriately define sources of uncertainty.

%  So many sources of interaction pooled into a single entity; hard to estimate.

% MLE is zero if no interaction is detected, but there may still be a substantial probability of interaction should the two species co-occur. [This is particularly important in the context of a changing world, where species not co-occurring today – or species not detected co-occurring today – may well do so tomorrow, forming complex interactions.]

@@

    %DG: note that I replaced theta by lambda to follow the above definitions. Need to make sure it's done throughout
    %DG: I realize there is an X in here that should be removed as well to avoid confusion with the above nomenclature

      We here adopt a Bayesian approach to estimate the distribution of the parameter $\lambda$. According to Bayes principle, the posterior distribution of $\lambda$ is:

      \begin{equation}
        \underbrace{P(\lambda|X,n)}_{Posterior} = \frac{\overbrace{P(X|\lambda,n)}^{Likelihood}\overbrace{P(\lambda)}^{Prior}}{\underbrace{P(X|N)}_{Normaliser}} .
        \label{posterior}
      \end{equation}

      According to the above description, the likelihood is simply the binomial distribution (Eq.~\ref{likelihood}). Since $\lambda$ is a probability it is bounded between 0 and 1 and consequently the most appropriate prior distribution is the beta:

      \begin{equation}
        \lambda \sim Beta(\alpha,\beta) , \label{prior}
      \end{equation}

@@

     In some cases it can be complicated to compute the normaliser, but fortunately in our case an analytical solution exists. The beta-binomial distribution is a conjugate distribution of the binomial distribution. This allows us to analytically compute the posterior distribution of a binomial model with a beta prior distribution. We can re-write the posterior distribution of $\lambda$ as:

      \begin{equation}
        P(\lambda|k,n) = \frac{\lambda^{\alpha+k-1}(1-\lambda)^{\beta+n-k-1}}{B(\alpha+k,\beta+n-k)} , \label{posterior}
      \end{equation}

%      \begin{equation}
%        Beta(\alpha+k,\beta+n-k) = \frac{\Gamma(\alpha+k)\Gamma(\beta+n-k)}{\Gamma(\alpha+\beta+n)} . \label{betafunction}
%      \end{equation}

      \noindent where the function $B$ is the beta function. The posterior distribution of $\lambda$ therefore follows the beta distribution with new parameters $\alpha'= \alpha+k$ and $\beta'=\beta+n-k$. The weigth of the prior on the posterior distribution is understood from these definition of the parameters: the difference between the posterior and the prior will increase with $k$ and $n-k$. When plotted, we find the shape of the distribution gets narrower with $k$ and $n$ (Fig.~\ref{Salix_pdfs}). We also find an important difference between the salix-gallers and gallers-parasitoids because of the differences in the prior distributions. 
      %
      %  The posterior distribution of $\theta$ could be computed with R with the following command:
      %
      %
      %  \vspace{12pt}
      %  \noindent\emph{dbeta(x = theta, shape1 = alpha+k, shape1 = beta+n-k)}
      %  \vspace{12pt}

%DG :  I think we should add the MLE estimate along with the binomial confidence interval on the figure in order to better illustrate the impact of the prior (obviously only for one N). You may draw it with a line to contrast with polygons.

%DG: otherwise the difference between the two interaction types is very interesting and useful to show the effect of the priors. Bravo ! 
      \begin{figure}[ht]
        \label{Salix_pdfs}
        \caption{Using prior distribution based on the \emph{Salix}-galler and galler-parasitoid networks in~\citet{Barbour2016}, we can calculate posterior distributions for the probability of interaction ($\lambda$) between two species that have not yet been observed interacting. Here we show posterior distributions for each $\lambda$ in each network ranging from the prior distribution (N=0 observed co-occurrence) to the distribution obtained when the pair of species has been observed co-occurring 100 times. The distribution narrows and approaches zero as the sample size increases. \textbf{A)} The posterior distributions for the \emph{Salix}-galler component are always wider and farther from zero than those for \textbf{B)} the galler-parasitoid component. This is likely because the prior distribution for the galler-parasitoid component of the network was both narrower and had a mean closer to that in the~\citet{Kopelke2017} data than the \emph{Salix}-galler component.}
        \includegraphics*[width=.8\textwidth]{figures/Salix_Galler_pdfs_increasing_N.eps}
        \end{figure}

@@

      The fact that the posterior distribution of $\lambda$ follows a beta distribution makes it straightforward to compute moments and other properties.

      The \textbf{average} of $\lambda$ is: 
          \begin{equation}
            \bar{\lambda} = \frac{\alpha+k}{\alpha+\beta+n} ,
            \label{mean}
          \end{equation}

        and its \textbf{variance} is:  
          \begin{equation}
            Var(\lambda|k) = \frac{(\alpha + k)(\beta + n - k)}{(\alpha + \beta + n)^{2}(\alpha + \beta + n +1)}
            \label{variance}
          \end{equation}

        The \textbf{mode} of the distribution is:
          \begin{equation}
            \hat{\lambda} = \frac{\alpha + k - 1}{\alpha + \beta + n - 2} .
            \label{mode}
          \end{equation}

@@

      Parameters $\alpha$ and $\beta$ determine the shape of the prior distribution, which follows a beta distribution. These are called hyper parameters. Below we identify four ways to formulate the prior distribution of $\lambda$.

@@

          In absence of any external information, an uninformative prior is the most conservative hypothesis for the distribution of $\lambda$. The beta distribution is in this case a uniform distribution, specified with hyper parameters $\alpha=1$ and $\beta=1$.

@@

          [[Should we still have this? At the workshop we talked about how interaction probabilities depend on the degrees of both species and so I'm using predicted interaction probabilities (degreeA*degreeB) in the galler example. A writeup for that process follows the original text:]]

%DG: I think so, it is a very useful proposition in the paper to increasingly improve the quality of the prior. I just tried to cut it down in order to follow the above layout. 
%DG: actually, once I got down to the following section, I realized that the example is based on the degree distribution, so there is not much choice to keep it

          The degree of a node in a network is defined as the number of interactions it has to other nodes. The degree distribution of a network is then the probability distribution of these degrees over the whole network and the standardised degree could therefore be interpreted as an interaction probability. It is consequently possible to use the degree distribution to inform the prior distribution. The degree distribution could come from several networks, from a similar network (e.g. a known network at slightly different location) or from the network of interest if interaction probabilities for some species are already documented. The latter approach allows us to apply information from known, abundant species to the rarest species for which interactions are less frequently documented. The procedure for the estimation of the hyper parameters follows the same approach as described above for connectance except that each measurement is at the individual species level instead of the network level.

          If our focal network describes a similar system to that in a known network, we can use the distribution of interaction probabilities in that network to inform our prior. The probability of any interaction $L_ij$ depends on the degrees of species $i$ and $j$. Using normalised degrees $\Delta_i$ and $\Delta_j$ (i.e., degrees divided by the number of species in the network), we can obtain the probability of interaction $L_ij$=$\Delta_i\times\Delta_j$. The distribution of these interaction probabilities derived from a similar network or a suite of networks can be used to establish a prior distribution before any data from the focal network are collected. We could also use this approach to build a prior distribution using known interaction probabilities for some species in the focal network. The latter approach allows us to apply information from known, abundant species to the rarest species for which interactions are less frequently documented. The procedure for the estimation of the hyper parameters follows the same approach as described above for connectance except that each measurement is at the individual interaction level instead of the network level.

@@

          [Tomas adds: This is a bit of a stump]
%DG: I tried to expand just a little, but I don't want to get lost in this paragraph as it would be a distraction. We should basically mention it as a possibility.
